Building a continuous benchmarking ecosystem in bioinformatics

### 2025-10-02 2_second_review/manuscript_with_tracked_changes.tex
Reorder eessi vs easybuild bibitems
Add Charlotte's and Mark's edits

{\color{red} Transparency and clear governance guidelines provide an effective means to detect and react to potential misuse, e.g., by explicit rules to validate results. Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency. Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide explicit means to ease collaboration. In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}

>>>>

{\color{red} Transparency and clear governance guidelines provide an effective means to detect and react to potential misuse, e.g., by explicit rules to validate results. Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency. Clear code of conduct statements and proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) provide explicit means to ease collaboration. In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}

### 2025-10-02 3_third_review/manuscript_with_tracked_changes.tex
Add Charlotte's and Mark's edits

Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single or} small number of researchers (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}. This brings some disadvantages. For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets {\color{red}without being aware that their colleagues are simultaneously undertaking similar efforts, leading to redundant and fragmented work}. Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results. However, because there are not many standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task. Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{luecken2025open}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.

{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} that is, welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this} centralization phase {\color{red}\sout{)}}. Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures. However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and requires additional measures that build trust (e.g., transparency).

>>>>

Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}, }being started by a {\color{red} single researcher or} small {\color{red}\sout{number of researchers}team} (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}. This brings some disadvantages. For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets {\color{red}without being aware that their colleagues are simultaneously undertaking similar efforts, leading to redundant and fragmented work}. Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results. However, because there are not many standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task. Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{luecken2025open}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.

{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} i.e., welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this the} centralization phase {\color{red}\sout{)}}. Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures. However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and requires additional measures that build trust (e.g., transparency).
